Estimating SARS-CoV-2 transmission in England from randomly sampled prevalence surveys

### 2024-02-13 transmission.tex

The probability of transmission, $q_{aa'}(t)$, needs further decomposition to be identifiable, for the same issue of dimensionality preventing each element of $\matr{\beta}$ being estimated independently (see \cref{SEIR:sec:structured-populations} for a full explanation).
I assume that $q_{aa'}(t) = \beta(t) q^\text{susc}_{a} q^\text{inf}_{a'}$ where $\beta(t)$ is a time-varying reference probability of transmission; $q^\text{susc}_a$ is the relative \emph{susceptibility} of stratum $a$, \ie how vulnerable to infection individuals in stratum $a$ are; and $q^\text{inf}_{a'}$ is the relative \emph{transmissibility} of stratum $a'$, \ie how likely individuals in stratum $a$ are to pass on an infection.
Separating relative transmissibility and susceptibility into two independent factors is a common approach\todo{cite anderson and may}.
These measures include both biological differences, \eg due to differences in immune systems, or behavioural, \eg if infants are less hygienic than adults.

@@

These patterns can be interpreted as an instance of the common bias---variance trade-off.
The SEIR model imposes stronger assumptions about the epidemic dynamics.
The assumptions reduces the variance of the estimates, including by making them smoother, but if the assumptions are incorrect, bias is introduced.
This trade-off based on the strength of assumptions has long been known to apply in the backcalulcation context~\autocite[e.g.][section 8.3]{brookmeyerBackcalculation}.
The SEIR model can be viewed as a backcalulation method where the transmisson part of the model is a smoothing method.

>>>>

The probability of transmission, $q_{aa'}(t)$, needs further decomposition to be identifiable, for the same issue of dimensionality preventing each element of $\matr{\beta}$ being estimated independently (see \cref{E-SEIR:sec:structured-populations} for a full explanation).
I assume that $q_{aa'}(t) = \beta(t) q^\text{susc}_{a} q^\text{inf}_{a'}$ where $\beta(t)$ is a time-varying reference probability of transmission; $q^\text{susc}_a$ is the relative \emph{susceptibility} of stratum $a$, \ie how vulnerable to infection individuals in stratum $a$ are; and $q^\text{inf}_{a'}$ is the relative \emph{transmissibility} of stratum $a'$, \ie how likely individuals in stratum $a$ are to pass on an infection.
Separating relative transmissibility and susceptibility into two independent factors is a common approach\todo{cite anderson and may}.
These measures include both biological differences, \eg due to differences in immune systems, or behavioural, \eg if infants are less hygienic than adults.

@@

These patterns can be interpreted as an instance of the common bias---variance trade-off.
\todo{Comment on SEIR not fully adjusted for non-response bias}
The SEIR model imposes stronger assumptions about the epidemic dynamics.
The assumptions reduces the variance of the estimates, including by making them smoother, but if the assumptions are incorrect, bias is introduced.
This trade-off based on the strength of assumptions has long been known to apply in the backcalulcation context~\autocite[e.g.][section 8.3]{brookmeyerBackcalculation}.
The SEIR model can be viewed as a backcalulation method where the transmisson part of the model is a smoothing method.
